StackTokens: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-03 fully-abstract-comp/popl2019/paper.tex
Added arrow to alert reader to the address space layout in figures.
Finished text for example for check base.

\section{Linear Stack and Return Capabilities}
\todo[inline]{The stack grows downwards, maybe we should make the figures reflect this. Also make sealed capabilities more clear by greying out.}
\begin{figure}
  \centering
  \begin{subfigure}{0.4\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}

@@

% Prevention of attack 2
In a traditional calling convention with a single stack, the stack is used as a call stack in the sense that it keeps track of the order of the calls.
If everyone follows the calling convention, then a call pushes a stack frame to the call stack and a return pops the topmost stack frame from the call stack.
This convention is not followed by the adversarial code in Figure~\ref{fig:ret-ptr-abuse} as the adversarial code skips two stack frames.
In the presence of adversarial code, we need some enforcement mechanism that forces the adversarial code to follow the call stack order.
One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
The callee would have to give up this token on return to make sure that the token is only spent once.
As it turns out, the stack capability for the unused part of the stack can be seen as such a token, so when the callee returns, we expect them to return the stack capability we gave them in the call.
When we receive a stack capability on return, we will try to splice it with the now unsealed stack capability for our private stack frame.
If this is successful, then we know that the two capabilities are adjacent.
If it is unsuccessful or if no stack capability is returned, then we cannot determine whether our stack frame is the topmost.
The issue illustrated in Figure~\ref{fig:ret-ptr-abuse} is prevented by splicing as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}.
\begin{figure}
  \centering
  \begin{subfigure}{0.4\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}

@@

% known stack base
% + more
We check that the returned stack capability is adjacent to the stack capability for our local stack frame, but this does not guarantee that it is the same token as we gave away in the call.
Specifically, we have no idea whether the callee has kept part of the token for themselves and only given us back the part that is adjacent to our token.
If we do not make sure that the entire token is returned, then it is possible for an adversary to break WBCF.
This is illustrated in Figure~\ref{fig:stk-base-abuse}: Some trusted code has created a stack frame with their local data and called a piece of untrusted code giving them access to the remainder of the stack.
This situation is illustrated in Figure~\ref{fig:stack-base-abuse-a}.
Now the adversarial code keeps the stack capability for the part of the stack adjacent to the trusted code's stack frame.
The adversary does not seal this part in a return pair, but simply stores it on the heap.
At this point the adversary calls the trusted code with the remainder of the stack capability as illustrated in Figure~\ref{fig:stack-base-abuse-b}.
At some point the trusted code calls the adversary again leaving a second stack frame on the stack an giving the adversary the rest.
At this point, the adversary has access to the stack capability passed in the second call performed by the trusted code as well as the stack capability stored away after the first call (Figure~\ref{fig:stack-base-abuse-c}).
Now the adversary has access to a stack capability for part of the stack adjacent to the first stack frame of the trusted caller.
The adversary can use this stack capability to return from the first call without ever returning from the second call (Figure~\ref{fig:stack-base-abuse-d}).

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.23\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \actadv{(0,4)}{(4.5,7)}{}

>>>>

\section{Linear Stack and Return Capabilities}
\todo[inline]{The stack grows downwards, maybe we should make the figures reflect this. Also make sealed capabilities more clear by greying out.}
\begin{figure}
  \centering
  \begin{subfigure}{0.4\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \draw (-0.5,13) node {\footnotesize 0};
      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}

@@

% Prevention of attack 2
In a traditional calling convention with a single stack, the stack is used as a call stack in the sense that it keeps track of the order of the calls.
If everyone follows the calling convention, then a call pushes a stack frame to the call stack and a return pops the topmost stack frame from the call stack.
This convention is not followed by the adversarial code in Figure~\ref{fig:ret-ptr-abuse} as the adversarial code skips two stack frames.
In the presence of adversarial code, we need some enforcement mechanism that forces the adversarial code to follow the call stack order.
One way to enforce this would be to hand out a token on call that can only be used when the caller's stack frame is on top of the call stack.
The callee would have to give up this token on return to make sure that the token is only spent once.
As it turns out, the stack capability for the unused part of the stack can be seen as such a token, so when the callee returns, we expect them to return the stack capability we gave them in the call.
When we receive a stack capability on return, we will try to splice it with the now unsealed stack capability for our private stack frame.
If this is successful, then we know that the two capabilities are adjacent.
If it is unsuccessful or if no stack capability is returned, then we cannot determine whether our stack frame is the topmost.
The issue illustrated in Figure~\ref{fig:ret-ptr-abuse} is prevented by splicing as illustrated in Figure~\ref{fig:ret-ptr-abuse-prev}.
\begin{figure}
  \centering
  \begin{subfigure}{0.4\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \draw (-0.5,13) node {\footnotesize 0};
      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \inactadv{(0,4)}{(4.5,7)} {\footnotesize Adv. stack frame 1}
      \inactsf{(0,7)}{(4.5,9)} {\footnotesize Trusted \\ \footnotesize stack frame 2}
      \actadv{(0,9)}{(4.5,12)} {\footnotesize Adv. stack frame 2}

@@

% known stack base
% + more
We check that the returned stack capability is adjacent to the stack capability for our local stack frame, but this does not guarantee that it is the same token as we gave away in the call.
Specifically, we have no idea whether the callee has kept part of the token for themselves and only given us back the part that is adjacent to our token.
If we do not make sure that the entire token is returned, then it is possible for an adversary to break WBCF.
This is illustrated in Figure~\ref{fig:stk-base-abuse}: Some trusted code has created a stack frame with their local data and called a piece of untrusted code giving them access to the remainder of the stack.
This situation is illustrated in Figure~\ref{fig:stack-base-abuse-a}.
Now the adversarial code keeps the stack capability for the part of the stack adjacent to the trusted code's stack frame.
The adversary does not seal this part in a return pair, but simply stores it on the heap.
At this point the adversary calls the trusted code with the remainder of the stack capability as illustrated in Figure~\ref{fig:stack-base-abuse-b}.
At some point the trusted code calls the adversary again leaving a second stack frame on the stack an giving the adversary the rest.
At this point, the adversary has access to the stack capability passed in the second call performed by the trusted code as well as the stack capability stored away after the first call (Figure~\ref{fig:stack-base-abuse-c}).
Now the adversary has access to a stack capability for part of the stack adjacent to the first stack frame of the trusted caller.
The adversary can use this stack capability to return from the first call without ever returning from the second call thus breaking WBCF (Figure~\ref{fig:stack-base-abuse-d}).
In this scenario, the issue arises because we do not check whether the entire token was returned to us.
We could do this by checking whether the base address\footnote{The stack grows downwards in the address space} of the token returned to us correspond to the base address of the token we provided in the call.
In the scenario we just sketched, we would be alerted to an attempt to break WBCF as the base address of the token returned to us in Figure~\ref{fig:stack-base-abuse-d} does not correspond to the base address of the token we passed to the adversary in Figure~\ref{fig:stack-base-abuse-a}.

@@

\begin{figure}
  \centering
  \begin{subfigure}{0.23\linewidth}
    \centering
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
      % recurrent parts
      \draw (-0.5,13) node {\footnotesize 0};
      \draw (-0.5,12.5) edge[thin,-latex] (-0.5,11);
      \stdstackstart[13]
      \inactsf{(0,2)}{(4.5,4)} {\footnotesize Trusted\\ \footnotesize stack frame 1}
      \actadv{(0,4)}{(4.5,7)}{}
